preciseTAD: A machine learning framework for precise TAD boundary prediction at base level resolution

### 2020-08-09 Manuscript/manuscript.Rmd
Adjust TAD conservation refs

TAD boundaries are known to be relatively conserved across cell lines [@dixon2012topological; @krefting2018evolutionary; @nora2012spatial; @dixon2015chromatin; @rao20143d; @Schmitt2016aa]. To assess conservation, we evaluated the overlap between cell line-specific boundaries detected by *preciseTAD*, Arrowhead, and Peakachu after flanking them by the resolution of the Hi-C data. Boundaries involving Arrowhead were flanked by 5 kb, while boundaries involving Peakachu were flanked by 10 kb. We found that 35% of *preciseTAD* predicted TAD boundaries and 55% of *preciseTAD* predicted loop boundaries were were conserved between cell lines, compared to only 26% and 55% for Arrowhead and Peakachu alone (Figure 5E). The better conservation of *preciseTAD*-predicted boundaries further support the notion of their higher biological relevance.

@@

Guided by both low-resolution Hi-C data and high-resolution genome annotation data, *preciseTAD* predicts base-level resolution boundaries, alleviating resolution limitations of Hi-C data. We show that, unlike Arrowhead, *preciseTAD* does not inflate the number of predicted boundaries, instead, providing only the most biologically meaningful boundaries that demarcate regions of high inter-chromosomal interactions. *preciseTAD* TAD and loop boundaries were enriched for known architectural transcription factors including CTCF, RAD21, SMC3, and ZNF143, supporting recent observations that TADs and loops are dynamic structures generated by similar mechanisms [@hansen2018recent]. Likewise, *preciseTAD* boundaries were found to be more conserved between GM12878 and K562 cell lines, a known feature among the 3D architecture of the human genome, further highlighting their biological significance [@dixon2012topological; @krefting2018evolutionary; @nora2012spatial; @dixon2015chromatin; @rao20143d; @Schmitt2016aa].

>>>>

TAD boundaries are known to be relatively conserved across cell lines [@dixon2012topological; @krefting2018evolutionary; @nora2012spatial; @sexton2012three]. To assess conservation, we evaluated the overlap between cell line-specific boundaries detected by *preciseTAD*, Arrowhead, and Peakachu after flanking them by the resolution of the Hi-C data. Boundaries involving Arrowhead were flanked by 5 kb, while boundaries involving Peakachu were flanked by 10 kb. We found that 35% of *preciseTAD* predicted TAD boundaries and 55% of *preciseTAD* predicted loop boundaries were were conserved between cell lines, compared to only 26% and 55% for Arrowhead and Peakachu alone (Figure 5E). The better conservation of *preciseTAD*-predicted boundaries further support the notion of their higher biological relevance.

@@

Guided by both low-resolution Hi-C data and high-resolution genome annotation data, *preciseTAD* predicts base-level resolution boundaries, alleviating resolution limitations of Hi-C data. We show that, unlike Arrowhead, *preciseTAD* does not inflate the number of predicted boundaries, instead, providing only the most biologically meaningful boundaries that demarcate regions of high inter-chromosomal interactions. *preciseTAD* TAD and loop boundaries were enriched for known architectural transcription factors including CTCF, RAD21, SMC3, and ZNF143, supporting recent observations that TADs and loops are dynamic structures generated by similar mechanisms [@hansen2018recent]. Likewise, *preciseTAD* boundaries were found to be more conserved between GM12878 and K562 cell lines, a known feature among the 3D architecture of the human genome, further highlighting their biological significance.
